---
title: "Assignment Practical Machine Learning"
author: "Jimmy Hipolito"
date: "16/12/2020"
output:
  html_document: default
  pdf_document: default
---



## Assignment: Prediction Assignment Writeup

by:  Jimmy Hipolito
date: december 2020


# 1. Overview

This proyect is about of quantify how well they do a particular activity. We will use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants.


# 2. Library

```{r}

library(caret) #version 6.0-86
library(knitr) #version 1.30
library(rpart) #version 4.1-15
library(rattle) #version 5.4.0
library(randomForest) #version 4.6-14
library(lattice) #version 0.20-41
library(ggplot2) #version 3.3.2

```


# 3. Datasets

The training and the testing data are available in the Course Project section of the Practical Machine Learning course.  After downloading the data, it is cleared of null values and dimensions are checked.

```{r}

trainUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testUrl  <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

#cleaning
train <- read.csv(url(trainUrl), na.strings=c("NA","#DIV/0!",""))
test <- read.csv(url(testUrl), na.strings=c("NA","#DIV/0!",""))

#checking dimension
dim(train)
dim(test)

```


# 4. Data Cleaning

In this section, the variables with missing values are eliminated, and finally the necessary variables are selected (the first 7 variables are removed, as they contain non-relevant information)

```{r}

train <-train[,colSums(is.na(train)) == 0]
test <-test[,colSums(is.na(test)) == 0]

train <-train[,-c(1:7)]
test <-test[,-c(1:7)]

dim(train)
dim(test)

```


# 5. Dataset Partitioning

The training dataset is partitioned in 2 sets: a Training set (70% of the data) and a Test set (30% of the data).  The testing dataset, composed of 20 cases, will only be used to resolve the final quiz.

```{r}

# Partitions training dataset
intrain <- createDataPartition(train$classe, p=0.7, list=FALSE)
training <- train[intrain, ]
testing <- train[-intrain, ]

dim(training)
dim(testing)

```
```{r}

barplot(table(training$classe), col="orange", main="Levels collected on 8 hours of activities", xlab="classe", ylab="Frequency")

```


# 6 Prediciton 

In this section, three methods learned in the course will be applied: Decision Trees, Random Forest and Generalized Boosted Model (GBM).  The model with the best Accuracy will be chosen for the final prediction with the test data. 


# a. Decision Trees

```{r}

set.seed(33833)

# Fitting model
model_dt <- rpart(classe ~ ., data=training, method="class")
# Predicting
predict_dt <- predict(model_dt, newdata=testing, type="class")
#Testing
confusionMatrix(table(predict_dt, testing$classe))

```

# b. Random Forest

```{r}

set.seed(33833)

# Fitting model
model_rf  <- train( classe ~.,data = training,method ="rf",trControl= trainControl(method="cv",number=3))
# Predicting
predict_rf <- predict(model_rf, newdata=testing)
# Testing
confusionMatrix(table(predict_rf, testing$classe))

```

# c. Generalized Boosted Model (GBM)

```{r}

set.seed(33833)

# Fitting model
model_gbm  <- train( classe ~.,data = training,method ="gbm",trControl= trainControl(method="repeatedcv",number=3,repeats=1),verbose=FALSE)
# Predicting
predict_gbm <- predict(model_gbm, newdata=testing)
# Testing
confusionMatrix(table(predict_gbm, testing$classe))


```


# 7. Prediction

The accuracy of the 3 regression modeling methods are:

Decision Trees:  74.1%
Random Forest:   99.2%
GBM:             96.1%

In conclusion, the Random Forest model will be applied to predict the 20 quiz results.

```{r}

predict_test <- predict(model_rf, newdata=test)
predict_test

```


